---
title: "Chris_1"
author: "Chris Shen"
date: "11/26/2021"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r load-libraries, warning = FALSE, message = FALSE}
library(tidyverse)
library(janitor)
library(here)
library(parsnip)
```

### Introduction

This part of the presentation focuses on studying two main topics:

- Effect of GDP (developing vs developed world) on immunisation rate

- Effect of immunisation against diseases (Hepatitis B, Polio & Diphtheria) on under 5 deaths

Data is wrangled then explored through visualisations and summary statistics, followed by some modelling from which conclusions are drawn.

#### Variables

Predictor variables: 

- Average immunisation against disease

- GDP

- Status of country

Outcome variables: 

- Under 5 deaths

#### Hypotheses

- An increase in GDP results in higher immunisation rates

- Developed countries tend to have higher immunisation rates

- An increase immunisation rates results in reduced deaths

### Data

Read data and changed variable names to snake_case using the `clean_names()` function from the `janitor` package.
```{r clean-data}
life_expectancy <- read_csv(here("data/Life Expectancy Data.csv")) %>%
  clean_names()
```

Created new variable `average_immunity` by taking the average of the three variables `hepatitis_b`, `polio`, `diphtheria`. Saved this as new data frame named `immune`.
```{r immune}
immune <- life_expectancy %>%
  select(country, year, status, gdp, population, under_five_deaths, hepatitis_b, polio, diphtheria) %>%
  mutate(average_immunity = (hepatitis_b + polio + diphtheria) / 3) %>%
  mutate(average_immunity = round(average_immunity, digits = 0)) %>%
  select(country, status, year, gdp, population, under_five_deaths, average_immunity)

immune
```

### Visualisations: GDP & average immunisation

#### Plot of average immunisation (%) against GDP

```{r}
immune %>%
  ggplot(aes(x = gdp, y = average_immunity)) +
  geom_jitter()
```

#### Grouped by developed & developing status

```{r}
immune %>%
  group_by(status) %>%
  ggplot(aes(x = gdp, y = average_immunity, colour = status)) +
  geom_jitter()
```

#### Facet

```{r}
immune %>%
  group_by(status) %>%
  ggplot(aes(x = gdp, y = average_immunity, colour = status)) +
  geom_jitter() +
  facet_wrap(. ~ status) +
  scale_x_continuous(breaks = c(50000, 125000))
```

#### Statistics

```{r}
immune %>%
  group_by(status) %>%
  filter(!is.na(gdp), !is.na(average_immunity)) %>%
  summarise(mean(gdp), min(average_immunity), mean(average_immunity), max(average_immunity))
```


### Visualisations: Average immunisation & under 5 deaths

#### Plot of under 5 deaths against average immunisation (%)

```{r av-immune}
av_immune <- immune %>%
  select(year, average_immunity, under_five_deaths)
```

This plot I cannot tell head or tails.
```{r plot-under-5-deaths-against-average-immunisation}
av_immune %>%
  ggplot(aes(x = average_immunity, y = under_five_deaths)) +
  geom_jitter() +
  labs(title = "Plot of under 5 deaths against average immunisation (%)",
       x = "Average immunity (%)",
       y = "Under 5 deaths (per 1000 of population)")
```

So I faceted the plot by year and the message is much clearer.
```{r plot-under-5-deaths-against-average-immunisation-facet}
av_immune %>%
  group_by(year) %>%
  ggplot(aes(x = average_immunity, y = under_five_deaths)) +
  geom_jitter() +
  facet_wrap(. ~ year) +
  labs(title = "Plot of under 5 deaths against average immunisation (%)",
       subtitle = "By year",
       x = "Average immunity (%)",
       y = "Under 5 deaths (per 1000 of population)")
```

#### Plot of under 5 deaths against average immunisation (%) for year 2008

This year is selected as it works best with the available tool set, namely linear regression.
```{r av-immune-filter}
av_immune <- av_immune %>%
  filter(year == 2008)
```

```{r plot-under-5-deaths-against-average-immunisation-2000}
av_immune %>%
  ggplot(aes(x = average_immunity, y = under_five_deaths)) +
  geom_jitter() +
  labs(title = "Plot of under 5 deaths against average immunisation (%)",
       subtitle = "For year 2008",
       x = "Average immunity (%)",
       y = "Under 5 deaths (per 1000 of population)")
```

### Modelling

```{r trendline-plot}
av_immune %>%
  ggplot(aes(x = average_immunity, y = under_five_deaths)) +
  geom_jitter() +
  geom_smooth(method = "lm") +
  labs(title = "Plot of under 5 deaths against average immunisation (%)",
       subtitle = "For year 2008, with modelling",
       x = "Average immunity (%)",
       y = "Under 5 deaths (per 1000 of population)")
```

```{r plot-regression}
av_immune_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(under_five_deaths ~ average_immunity, data = av_immune)

tidy(av_immune_fit)
```

### Conclusion

Confirm hypothesis