---
title: "Chris"
author: "Chris Shen"
date: "11/25/2021"
output: 
  html_document: 
    keep_md: yes
editor_options: 
  chunk_output_type: console
---

```{r warning = FALSE, message = FALSE}
library(tidyverse)
library(janitor)
library(here)
library(parsnip)
```

## Data

```{r message = FALSE}
life_expectancy <- read_csv(here("data/Life Expectancy Data.csv")) %>%
  clean_names()
```

```{r}
immune <- life_expectancy %>%
  select(country, year, status, gdp, population, under_five_deaths, hepatitis_b, polio, diphtheria) %>%
  mutate(average_immunity = (hepatitis_b + polio + diphtheria) / 3) %>%
  mutate(average_immunity = round(average_immunity, digits = 0)) %>%
  select(country, status, year, gdp, population, under_five_deaths, average_immunity)

immune
```

## Visualisations

### A developing country: Afghanistan

```{r data}
immune_afg <- immune %>%
  select(country, year, gdp, population, under_five_deaths, average_immunity) %>%
  filter(country == "Afghanistan")
```

#### Plot 1: GDP over time

```{r gdp over time}
immune_afg %>%
  ggplot(aes(x = year, y = gdp)) +
  geom_col() +
  geom_smooth(method = "lm") +
  labs(title = "GDP of Afghanistan over time")
```

```{r model-gdp}
afg_gdp_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(gdp ~ year, data = immune_afg)

tidy(afg_gdp_fit)
```

#### Plot 2: Average immunisation over time

```{r}
immune_afg %>%
  ggplot(aes(x = year, y = average_immunity)) +
  geom_col() +
  labs(title = "Average immunisation (%) over time")
```

#### Plot 3: Population over time

```{r}
immune_afg %>%
  ggplot(aes(x = year, y = population)) +
  geom_col() +
  geom_smooth(method = "loess") +
  labs(title = "Population over time")
```

## Modelling 1 - Average immunisation (%) effect on under 5 deaths (per 1000)

#### Visualisations of under 5 deaths against average immunisation

```{r linear}
av_immune <- immune %>%
  select(average_immunity, under_five_deaths)

av_immune %>%
  ggplot(aes(x = average_immunity, y = under_five_deaths)) +
  geom_jitter() +
  geom_smooth(method = "lm") + 
  labs(title = "Linear model")
```

There is too much noise - let's do this for a specific year.

#### Visualisation & model for year 2015

```{r}
immune_2015 <- immune %>%
  filter(year == 2015) %>%
  select(average_immunity, under_five_deaths)
```

```{r}
immune_2015 %>%
  ggplot(aes(x = average_immunity, y = under_five_deaths)) +
  geom_jitter() +
  geom_smooth(method = "lm") + 
  labs(title = "Linear model")

immune_2015_fit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(under_five_deaths ~ average_immunity, data = immune_2015)

tidy(immune_2015_fit)
```

## Modelling 2 - GDP effect on Average immunisation (%) for year 2015

```{r}
immune_gdp <- immune %>%
  filter(year == 2015) %>%
  select(gdp, average_immunity)
```

```{r}
immune_gdp %>%
  ggplot(aes(x = gdp, y = average_immunity)) +
  geom_jitter() +
  geom_smooth()
```

Unclear what is happening - split into GDP intervals

#### GDP from 0 to 20000

```{r}
developing <- immune_gdp %>%
  filter(!gdp > 20000)
```

```{r}
developing %>%
  ggplot(aes(x = gdp, average_immunity)) +
  geom_jitter() +
  geom_smooth(method = "lm")
```

#### GDP over 20000

```{r}
developing <- immune_gdp %>%
  filter(gdp > 20000)
```

```{r}
developing %>%
  ggplot(aes(x = gdp, average_immunity)) +
  geom_jitter() +
  geom_smooth(method = "lm")
```
